---
title: "ACST8083 Assignment 1,"
author: "Ngoc Tuyet Trinh"
date: "2024-05-02"
output: html_document
---
Question 1

First, I load the data:
```{r}
set.seed(10)
Assg <- read.csv("Assignment1_Dataset_2024.csv")
```

The dataset has 607697 policies. There are 8 columns: Counts, exposure, x1 (weight), x2 (distance), x3 (age), x4 (carage), x5 (gender) and x6 (state). The summary() provides some statiscal measures of the numerical variables.
```{r}
str(Assg)
```
```{r}
summary(Assg)
```
There are no missing values in the dataset.
```{r}
colSums(is.na(Assg))
```
Check the mode of categorical variables:
```{r}
names(sort(-table(Assg$gender)))[1]
```
```{r}
names(sort(-table(Assg$state)))[1]
```
Change the categorical variables into type of factor
```{r}
Assg$gender<-as.factor(Assg$gender)
Assg$state<-as.factor(Assg$state)
```

Plot the numerical variables
```{r}
par(mfrow = c(2,2))
#Plotting Numerical data
hist(Assg$distance)
hist(Assg$weight)
hist(Assg$age)
hist(Assg$carage)
```

Plot the categorical variables
```{r}
gender_counts <- table(Assg$gender)
pie(gender_counts, main = "Gender Distribution")
```
```{r}
state_counts <- table(Assg$state)
pie(state_counts, main = "State Distribution")
```

Question 2

First, I fit the GLM model including x1,x2,x3,x4,x5,x6 and apply the anova() with chisq test to test the significance of the attributes:
```{r}
begin_model <-glm(Counts ~ weight + distance + age + carage + gender + state, data=Assg, 
                  family=poisson(), offset=log(exposure))
anova(begin_model,test ="Chisq")
```
Since x6 (state) is insignificant, I drop it. Then Model1's coefficients are:
```{r}
Model1 <-glm(Counts ~ weight + distance + age + carage + gender, data=Assg, 
             family=poisson(), offset=log(exposure))
Model1$coefficients
```
Let x1 = 2000, x2 = 15, x3 = 30, x4 = 4, x5 = female (don't use x6 because it is dropped already), calculate 位(x) :
```{r}
lambda<-predict(Model1, list(weight = 2000,
                                 distance = 15,
                                 age = 30,
                                 carage = 4,
                                 gender= "female",
                                 exposure = 1), type="response")
                                
lambda
```
Plot 位(x) versus x3 (age)
```{r}
x_age<-seq(min(Assg$age),max(Assg$age),1)
y_lambda<-predict(Model1, list(weight=rep(2000, length(x_age)),
                                      distance=rep(15,length(x_age)),
                                      age=x_age,
                                      carage=rep(4,length(x_age)),
                                      gender=rep("female",length(x_age)),
                                      exposure=rep(1,length(x_age))), type="response")

plot(x_age,y_lambda, main ="Intensity vs age", xlab="age",ylab="Intensity",col="green",lwd = 2, type="l")
```

Question 3

Fit the GLM model, use step() to do backward selection, then assign this model as Model2:
```{r}
start_model <- glm(Counts ~ weight + distance + age + carage + gender + state +
                     I(weight^2) + I(distance^2) + I(age^2) + I(carage^2),
                   data = Assg, family = poisson(), offset = log(exposure))
Model2<- step(start_model, direction ="backward", trace = 0)
Model2
```
Calculate 位(x) with x1 = 2000, x2 = 15, x3 = 30, x4 = 4, x5 = female
```{r}
lambda2<-predict(Model2, list(weight = 2000,
                             distance = 15,
                             age = 30,
                             carage = 4,
                             gender= "female",
                             exposure = 1), type="response")

lambda2
```
Plot 位(x) versus x3 (age)
```{r}
y_lambda2<-predict(Model2, list(weight=rep(2000, length(x_age)),
                               distance=rep(15,length(x_age)),
                               age=x_age,
                               carage=rep(4,length(x_age)),
                               gender=rep("female",length(x_age)),
                               exposure=rep(1,length(x_age))), type="response")

plot(x_age,y_lambda, main ="Intensity vs age", xlab="age",ylab="Intensity",col="green",lwd = 2, type="l")
lines(x_age,y_lambda2, xlab="age",ylab="Intensity",lwd = 2.5,col="red")
legend(x = "topright",  
       legend=c("Model1", "Model2"),  
       fill = c("green","red") )
```


Question 4

The likelihood ratio test for Model1 and Model2:
```{r}
library(lmtest)
lrtest(Model1, Model2)
```
Question 5

10-fold cross validation error for Model1:
```{r}
library(boot)
#Set up the cost function
cost<-function(obs, pred){
  2*mean(pred - obs + log((obs/pred)^obs ))
} 
#Cross validation of GLM for Model1
CVerror_1<-cv.glm(Assg, Model1, cost, K=10)
#CVerror for Model1
CVerror_1$delta[1]
```
10-fold cross validation error for Model2:
```{r}
CVerr_2<-cv.glm(Assg, Model2, cost, K=10)
#CVerror for Model2
CVerr_2$delta[1]
```